---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BRINDA

<!-- badges: start -->
<!-- badges: end -->

The BRINDA R package is a user-friendly all-in-one R package that uses a series 
of functions to implement BRINDA adjustment [method](https://brinda-nutrition.org/publications/) .

## Installation

You can install the development version of BRINDA from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hanqiluo/BRINDA")


```

## Example

This is a basic example which shows you how to adjust retinol binding protein, 
    retinol, ferritin, soluble transferrin receptor, and zinc using AGP and/or 
    CRP.

```{r example}
library(BRINDA)
## basic example code
data(sample_data)

sample_data_adj <- BRINDA(dataset = sample_data,
       retinol_binding_protein_varname = rbp,
       retinol_varname = sr, ferritin_varname = sf,
       soluble_transferrin_receptor_varname = stfr,
       zinc_varname = zinc, crp_varname = crp,
       agp_varname = agp, population = Psc,
       crp_ref_value_manual = ,
       agp_ref_value_manual = ,
       output_format = )

```

## Citation
Luo, H.;Addo, Y.;Jahan, A (2021) BRINDA: Computation of BRINDA Adjusted Micronutrient Biomarkers for inflammation. R package version 0.1.2, https://github.com/hanqiluo/BRINDA

## Contributing
If you would like to report bugs, suggest features, or leave comments, please [create issues](https://github.com/hanqiluo/BRINDA/issues).

If you would like to contribute code, please fork the source code, modify, and issue a [pull request](https://github.com/hanqiluo/BRINDA/pulls).

## Acknowledgements
The Authors thank all members of the BRINDA working group who helped in developing the BRINDA adjustment method. The authors also thank Charles D. Arnold, Fanny Sandalinas, Kevin Tang, and Lucas Gosdin for the extensive testing of the package and thank Joanne Arsenault for her editing and comments.
